---
title: "Transform Data"
---


# Import Data

## Your Turn 1

Set the working directory to `03-Transform`. Then import the `pbc.csv` data set, which is in the `03-Transform` folder, as `pbc`. Then copy the import code into the code chunk below. Does it run?

```{r}
library(tidyverse)
pbc <- read_csv("pbc.csv")
```

```{r}
pbc
```


# Transforming Data with `dplyr`

#### Part 1, `select()`, `filter()`, `arrange()`

**SELECT**
Select the id column from pbc
```{r}
select(pbc, id)
```
The above code is the same as
```{r}
pbc %>% 
  select(id)
```

You can use helpers with select, for example to select everything but the id and age
```{r}
pbc %>% 
  select(-c(id,age))
```
You can select every column that mentions time with:
```{r}
pbc %>% 
  select(contains('time'))
```

Use the data-transformation-cheatsheet to find the helper that would let you select every column that starts with p
```{r}
pbc %>% 
  select(starts_with('p'))
```


**FILTER**
You can filter for patients who received the placebo
```{r}
pbc %>% 
  filter(trt == 'placebo')
```
You can filter for patients between age 30 and 60
```{r}
pbc %>% 
  filter(age>30, age<60)
```

We can exclude patients that got a transplant:
```{r}
pbc_filtered<-pbc %>% 
  filter(status !=1) 
pbc_filtered
```
Women between age 30 and 60 are at high risk of pbc, can you filter the data to create a high_risk subset?
```{r}
high_risk <- pbc %>% filter(
  age>30, age<60, sex == 'f'
) 
```

Could you add `arrange` to the above code so the data are sorted by `alk.phos` value.

#### Part 2: `mutate()`, `group_by()`, `summarize()`

```{r}
bmi <- tibble(
  plan = c('a','a','b','b','a'),
  weight_kg=c(96,87,104,61,131), 
  height_cm=c(174,189,195,149,131))

bmi <- bmi %>% mutate(bmi = weight_kg / ((height_cm)^2*.0001))

```


```{r}
bmi %>% group_by(plan) %>% mutate(group_avg_height= mean(height_cm))
```

```{r}
bmi %>% group_by(plan) %>% summarize(mean_bmi = mean(bmi), count= n())
```
*Practice*

Take the `pbc` data set and group by stage, then summarize to get the average `alk.phos` value, number of observations, and 
```{r}

```
You might see mean_alk.phos is filled with NA values, that's because we had some missing values in our `pbc` dataset.  By default the `mean()` returns NA if the dataset has any missing values. We can override this by setting the argument `na.rm = TRUE` this tells the mean funciton to ignore missing values can compute mean based on whatever is present. 

Change the code below so that we have a summary table of the number of patients in each stage for each sex. What variables are we grouping by and which are we summarizing? 

```{r}
pbc %>% 
  group_by(stage, sex) %>% 
  summarize(
    patient_count = n()
  )
```

Open the data-transformations-cheatsheet, look at the summary functions. Try adding another summary function to the summary dataframe you created above. 

### A few other handy dplyr tricks

Case when lets you conditionally apply something, Here we conditionally assign the new spiders column present or absent based on coding in the original spiders column
```{r}
pbc <- pbc %>% mutate(
  spiders = case_when(
    spiders == 1 ~ "present",
    spiders == 0 ~ "absent"
  )
)
```

Count the number of observations, even the number of observations by groups 

```{r}
pbc %>% 
  count(sex)
```

```{r}
pbc %>% 
  count(sex,trt,stage)
```

***

# Take aways

* Extract variables with `select()`  
* Extract cases with `filter()`  
* Arrange cases, with `arrange()`  

* Make tables of summaries with `summarise()`  
* Make new variables, with `mutate()`  
* Do groupwise operations with `group_by()`

* Connect operations with `%>%`  


